Metadata and Other Structures in texor

### 2025-09-07 review/response_to_review.Rmd

The packages texor and rebib have clear workflows, and their documentation are well-written.

I didn't fully review the source code since it was rather lengthy, but I could see room for improvement, e.g.,

- The stringr wrappers in texor are not necessary: https://github.com/Abhi-1U/texor/blob/e669ec3f9bf7ad5fb5c4b929f13d18ad7944f735/R/string-tools.R#L3-L28 What we need is `@importFrom stringr str_trim str_detect ...`, and the stringr functions can be used without the `stringr::` qualifier.

*Done, will be included in the latest CRAN release of the package.*

@@

    i <- grep(pattern, raw_lines) 
    raw_lines[i] <- gsub(target, replacement, raw_lines[i])

*Done, will be included in the latest CRAN release of the package.*

- The supplementary materials (https://github.com/Abhi-1U/texor-rjarticle/tree/master/supplement) are quite helpful, and I really like the examples. It will be nice to also include the intermediate `.Rmd` or `.md` files when converting `.tex` to `.html`, so readers can see how various LaTeX environments are converted to their Markdown counterparts, although the article has already provided some examples in Section 2. The function `texor::latex_to_web()` doesn't seem to have an argument for preserving the intermediate Markdown output, and it may be worth adding one.

*Done, intermediate `.Rmd` and `.md` files would be included in the next iteration of article. They are generated via the `texor::latex_to_web()` function and, by default, are retained automatically after a successful conversion without requiring explicit arguments.*

### 2025-09-07 supplement/figure-env/RJwrapper.Rmd
Final response for inclusion of Rmd files.

---
title: Working with figure environments in texor
abstract: This is a small sample article to demonstrate usage of texor to convert
  figure environments.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2009-06-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 1
issue: 1
slug: figure-env
packages:
  cran: texor
  bioc: []
preview: preview.png
bibliography: figure-example.bib
CTV: ~
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    legacy_pdf: yes
    web_only: yes
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

@@

::: article
# Introduction

Images are an essential component in any article, However due to the
differences in the support for various graphic formats between LaTeX and
markdown/HTML we need to fallback on raster graphics. It is well
summarized in the table [1](#table:1)

@@

  : Table 1: Image Format support in various Markup/Typesetting
  Languages
:::

# Image with width parameters

This section may contain a figure such as Figure [1](#figure:rlogo).
This is the most basic example of figure.

@@

![Figure 1: The logo of R.](Rlogo-5.png){#figure:rlogo width="35.0%"}

# Images in PDF format

Image [2](#fig:normal) is a graphical representation of normal
distribution.

``` r
\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\textwidth]{normal}
  \caption{PDF of a normal distribution}
  \label{fig:normal}
\end{figure}
```

![Figure 2: PDF of a normal distribution](normal.png){#fig:normal
width="50.0%"}

# Multiple images

Pandoc v3 and above now supports a new Figure object [@pandoc] which
supports multiple images side by side or in a grid format.

### Two or more Images side by side

@@

<figure id="fig:twoimages">
<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
src="normal.png" style="width:45.0%" alt="image" /></p>
<figcaption>Figure 3: Images side by side</figcaption>
</figure>

### Four Images in a grid

@@

<figure id="fig:fourimages">
<p><img src="Rlogo-5.png" style="width:45.0%" alt="image" /><img
src="normal.png" style="width:45.0%" alt="image" /></p>
<p><img src="normal.png" style="width:45.0%" alt="image" /><img
src="Rlogo-5.png" style="width:45.0%" alt="image" /></p>
<figcaption>Figure 4: Multiple images in a grid</figcaption>
</figure>

# Tikz images

Here is a tikz image example in fig [5](#fig:tikz) adapted from
[@casflow].

Another interesting aspect of including tikz image here is that you can
modify the source and re-convert without making any other changes.
Figure will get updated in the generated html article.

### Tikz Code:

``` r
\begin{figure}
\includegraphics{tikz/figtikz.png}

%% Generated Image will included as a PNG above automatically
  \centering
\tikzstyle{process} = [rectangle, rounded corners,
minimum width=3cm, 
minimum height=1cm,
text centered, 
draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]
\begin{tikzpicture}[node distance=4cm]
%Nodes
\node (start) [process] {LaTeX Source with Tikz Image};
\node (isolate) [process, below right of=start] {Isolated tikz env};
\node (pro1) [process, below left of=isolate] {Standalone PDF file};
\node (pro2) [process, below right of=pro1, yshift=-0.5cm] {PNG file};
\node (link) [process, below left of=pro2] {Include PNG in article};
\node (stop) [process, below right of=link] {Convert To RJ-web-article};
% arrows
\draw [arrow] (start) -- node[anchor=east] {Extract Figure Block} (isolate);
\draw [arrow] (isolate) -- node[anchor=east] {Compile} (pro1);
\draw [arrow] (pro1) -- node[anchor=east] {Convert to raster} (pro2);
\draw [arrow] (pro2) -- node[anchor=east] {Include back to source as image} (link);
\draw [arrow] (link) -- node[anchor=east] {Using Pandoc} (stop);
\end{tikzpicture}
\caption{Tikz Image example}
  \label{fig:tikz}
\end{figure}
```

### Resultant Figure :

<figure id="fig:tikz">
<img src="tikz/figtikz.png" style="width:100.0%" />

<figcaption>Figure 5: Tikz Image example</figcaption>
</figure>

# Algorithm2e diagrams

we do support algorithm2e diagrams and images, these will be numbered
differently and we strongly suggest to use \"alg:\" in labels for best
results. Here is an example of algorithm [6](#alg:how), referenced from
[@algoexample]

``` r
\begin{figure}[htbp]
\includegraphics{alg/alghow.png}

\SetAlgoLined
\KwData{this text}
\KwResult{how to write algorithm with \LaTeX2e }
initialization\;
\While{not at end of this document}{
read current\;
\eIf{understand}{
go to next section\;
current section becomes this one\;
}{
go back to the beginning of current section\;
}
}
\caption{How to write algorithms}
  \label{alg:how}
\end{figure}
```

<figure id="alg:how">
<img src="alg/alghow.png" style="width:100.0%" />
<p>initialization</p>
<figcaption>Algorithm 1: How to write algorithms</figcaption>
</figure>

# Summary

In summary the [*texor*](https://CRAN.R-project.org/package=texor)
package supports:

-   Almost all image formats in LaTeX.

-   Algorithm and tikz as well in some capacity.

@@

-   Image Captions with Numbering and Labelling.
:::

>>>>

---
title: Working with figure environments in texor
abstract: |
  This is a small sample article to demonstrate usage of
  [**texor**](https://CRAN.R-project.org/package=texor) to convert
  figure environments.
author:
- name: Abhishek Ulayil
  affiliation: Student, Institute of Actuaries of India
  orcid: 0009-0000-6935-8690
  address: Mumbai, India
date: '2025-09-01'
date_received: ~
journal:
  firstpage: ~
  lastpage: ~
volume: 17
issue: 3
slug: RJ-2025-000
citation_url: https://rjournal.github.io/
packages:
  cran: texor
  bioc: []
preview: preview.png
bibliography: figure-example.bib
CTV: ~
legacy_pdf: yes
legacy_converted: yes
output:
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
    mathjax: https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-chtml-full.js
    md_extension: -tex_math_single_backslash

@@

:::: article
## Introduction

Images are an essential component of any article, however, due to the
differences in support for various graphic formats between LaTeX and
markdown/HTML we need to fall back on raster graphics. The support for
different image formats across markup languages is summarized in
Table [1](#tab:T1){reference-type="ref" reference="table:1"}.

@@

  : (#tab:T1) Image Format support in various Markup/Typesetting
  Languages
:::

## Image with width parameters

The following code includes an image with width parameters, producing
the output in Figure [1](#figure:rlogo){reference-type="ref"
reference="figure:rlogo"}.

@@

![Figure 1: The logo of R.](Rlogo-5.png){#figure:rlogo width="35.0%"
alt="graphic without alt text"}

This is the most basic example of figure. Similarly PDF figures will be
included as rasterized images.

## Multiple images

Pandoc v3 and above now support a new `Figure` object [@pandoc] which
supports multiple images side by side or in a grid format. The following
subsections demonstrate these capabilities.

#### Two or more Images side by side

@@

<figure id="fig:twoimages">
<p><img src="Rlogo-5.png" style="width:45.0%"
alt="graphic without alt text" /><img src="normal.png"
style="width:45.0%" alt="graphic without alt text" /></p>
<figcaption>Figure 2: Images side by side</figcaption>
</figure>

#### Four Images in a grid

@@

<figure id="fig:fourimages">
<p><img src="Rlogo-5.png" style="width:45.0%"
alt="graphic without alt text" /><img src="normal.png"
style="width:45.0%" alt="graphic without alt text" /></p>
<p><img src="normal.png" style="width:45.0%"
alt="graphic without alt text" /><img src="Rlogo-5.png"
style="width:45.0%" alt="graphic without alt text" /></p>
<figcaption>Figure 3: Multiple images in a grid</figcaption>
</figure>

## Tikz images

The **texor** package supports tikz images by rasterizing it and making
it web friendly. Figure [4](#fig:tikz){reference-type="ref"
reference="fig:tikz"} shows a tikz image adapted from [@casflow].

#### Tikz Code:

The image in Figure [4](#fig:tikz){reference-type="ref"
reference="fig:tikz"} is a graphical representation of how **texor**
handles tikz images:

-   Figures containing tikz images are isolated.

-   Tikz libraries are fetched from the wrapper file.

-   The Tikz code section is isolated into a standalone LaTeX file and
    compiled.

-   The compiled LaTeX file generates a PDF Image.

-   This is converted to PNG format for embedding in the HTML output.

-   `\includegraphics{tikz/somefigure.png}` is added to the figure
    environment.

-   A lua filter removes redundant text the figure environment.

If you use **texor** to convert your articles using
`texor::latex_to_web()` with `temp_mode=TRUE`(it is `TRUE` by default).
The resultant Rmarkdown/HTML file will not modify the contents of your
LaTeX file. In this case you can keep reloading the article after making
changes to the tikz images, without having to do the above steps
manually in case you are converting the article by hand.

<figure id="fig:tikz">
<img src="tikz/figtikz.png" style="width:100.0%"
alt="graphic without alt text" />

<figcaption>Figure 4: Tikz Image example</figcaption>
</figure>

## Algorithm2e diagrams

Diagrams and images using the `algorithm2e` environment are supported,
these will be numbered differently. We strongly suggest to use \"alg:\"
in labels for best results. As a part of the filtering requires \"alg:\"
to number algorithm references seperately, otherwise they will share
references with normal figures.

Algorithm [1](#alg:how){reference-type="ref" reference="alg:how"} is an
example from the **algorithm2e** vignette [@algoexample].

<figure id="alg:how">
<img src="alg/alghow.png" style="width:100.0%"
alt="graphic without alt text" />

<figcaption>Algorithm 1: How to write algorithms</figcaption>
</figure>

## Other elements in figure objects

Figures can also house non-image environments like code blocks. Code
blocks in figure environments would share numbering with normal figures,
similar to LaTeX. Here Figure [5](#code:example){reference-type="ref"
reference="code:example"} refers to a code block.

<figure id="code:example">
<div class="center">
<div class="sourceCode" id="cb1"><pre class="sourceCode r"><code class="sourceCode r"><span id="cb1-1"><a href="#cb1-1" aria-hidden="true" tabindex="-1"></a>code_in_figure <span class="ot">&lt;-</span> <span class="cf">function</span>() {</span>
<span id="cb1-2"><a href="#cb1-2" aria-hidden="true" tabindex="-1"></a>  <span class="cf">if</span> (pandoc_version <span class="sc">&gt;=</span> <span class="dv">3</span>) {</span>
<span id="cb1-3"><a href="#cb1-3" aria-hidden="true" tabindex="-1"></a>    <span class="fu">print</span>(<span class="st">&quot;code in figure supported&quot;</span>)</span>
<span id="cb1-4"><a href="#cb1-4" aria-hidden="true" tabindex="-1"></a>  }</span>
<span id="cb1-5"><a href="#cb1-5" aria-hidden="true" tabindex="-1"></a>  <span class="cf">else</span> {</span>
<span id="cb1-6"><a href="#cb1-6" aria-hidden="true" tabindex="-1"></a>    <span class="fu">print</span>(<span class="st">&quot;code in figure not supported&quot;</span>)</span>
<span id="cb1-7"><a href="#cb1-7" aria-hidden="true" tabindex="-1"></a>  }</span>
<span id="cb1-8"><a href="#cb1-8" aria-hidden="true" tabindex="-1"></a>}</span></code></pre></div>
</div>
<figcaption>Figure 5: Example Code inside Figure
environment</figcaption>
</figure>

## Limitations

The limitations of [**texor**](https://CRAN.R-project.org/package=texor)
package in figure handling are:

-   Animations using the **animation** package in LaTeX wont be
    supported. The authors can include GiF equivalents of the same in
    the Rmarkdown/web article.

-   Multi-page PDF files are currently not supported.

-   Usage of subfigure,sclaebox,wrapfigure environments will not work as
    intended.

-   Currently only `[width=X.X\textwdith]` format is supported for
    defining the width in web article (default width would be 100% in
    other cases).

-   Other `\inclduegraphics[..]` options will be ignored.

## Summary

In summary the [**texor**](https://CRAN.R-project.org/package=texor)
package supports:

@@

-   Image Captions with Numbering and Labelling.

-   Algorithm2e and tikz figures.
::::

### 2025-09-07 supplement/metadata/metadata.tex
Final response for inclusion of Rmd files.

Footnotes do not work in \verb|tabular| environments in LaTeX but do work in R Markdown tables\footnote{Footnotes in paragraph text work in both LaTeX and R markdown}. Custom footnote numbering works in LaTeX but does not work in R Markdown\footnote[21]{Forever 21 :)}. These features are demonstrated in Table~\ref{table:3}.

\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\acronym{A}\footnote{This works in HTML}| & \acronym{A}\footnote{This works in HTML} \\ \hline
\verb|\R\footnote[21]{HTML shows 2 rather than 21}| & \R\footnote[21]{HTML shows 2 rather than 21}  \\ \hline
\end{tabular}
\caption{Demonstration of footnotes.}
\label{table:3}
\end{table}

@@

\paragraph{Note}
Certain reference commands like \verb|\nameref{}| and \verb|\autoref{}| will not work as intended in HTML.

>>>>

Footnotes do not work in \verb|tabular| environments in LaTeX but do work in R Markdown tables\footnote{Footnotes in paragraph text work in both LaTeX and R markdown}. Custom footnote numbering works in LaTeX but does not work in R Markdown\footnote[25]{Footnote 25 }. These features are demonstrated in Table~\ref{table:3}.

\begin{table}[htbp]
\centering
\begin{tabular}{| c | c |}
 \hline
 Command & Render \\
 \hline
 \verb|\acronym{A}\footnote{This works in HTML, but not in LaTeX}| & \acronym{A}\footnote{This works in HTML, but not in LaTeX}} \\ \hline
\verb|\R\footnote[25]{HTML shows 4 rather than 25}| & \R\footnote[25]{HTML shows 4 rather than 25}  \\ \hline
\end{tabular}
\caption{Demonstration of footnotes.}
\label{table:3}
\end{table}

@@

\paragraph{Note}
Certain reference commands like \verb|\nameref{}| and \verb|\autoref{}| may not work as intended in HTML.
